Multi-target detection with application to cryo-electron microscopy

### 2019-02-28 inverse_problems/inverse_problems.tex
./

In this section, we give the expressions for the autocorrelations of the observed signal $y$ under the Poisson model. In this model, each location $k$, $1 \le k \le N-L+1$, contains a Poisson-distributed number of random signals. More precisely, lettin $M_k$ be the number of signals starting at location $k$, $M_k \overset{iid}{\sim} \Poisson(\gamma / L)$. Note that the expected total number of signals occurring in $y$ is equal to $(N - L + 1)\gamma / L$, and consequently the total lengths of all signals (including overlaps) divided by $N$ is equal to
%
\begin{align}
\frac{N - L + 1}{N} \cdot L \cdot \frac{\gamma  }{L} 
\quad \overset{N\to\infty}{\longrightarrow} \quad \gamma.
\end{align}
%
Therefore, in the large $N$ limit $\gamma$ may be interpreted as the signal density, as in the well-separated model.

We will show the following expressions for the autocorrelations of $y$, expressed in terms of the autocorrelations of $x$ and the parameters $\gamma$ and $\sigma$:
%
\begin{align}
a_y^1  = & \gamma a_{x}^1, 
\label{eq:mean_micrograph2} \\
a_y^2[\ell] = & (\gamma a_x^1)^2 + \gamma a_x^2[\ell] + \sigma^2 \delta[\ell], 
\label{eq:ac2_micrograph2} \\
a_y^3[\ell_1,\ell_2] = & (\gamma a_x^1)^3 + \gamma a_x^1  
\cdot ( \gamma a_x^2[\ell_1] + \gamma a_x^2[\ell_2]
+ \gamma a_x^2[\ell_2-\ell_1]) + \gamma a_x^3[\ell_1,\ell_2] 
\nonumber \\
& + \gamma a_x^1 \sigma^2 \big(\delta[\ell_1]+\delta[\ell_2] +\delta[\ell_1-\ell_2] \big).
\label{eq:ac3_micrograph2}
\end{align}

@@

It is now easy to show how to solve for both $\gamma$ and $\sigma$ from these expressions, assuming $x$ is generic. Indeed, we observe the product
%
\begin{align}
(\gamma a_x^1) (\gamma a_x^2[1])
&=  \frac{\gamma^2}{L^2} \left( \sum_{i=0}^{L-1} x[i] \right)
\left(\sum_{j=0}^{L-1} x[j]x[j+1] \right)
\nonumber \\
&= \frac{\gamma^2}{L^2} \sum_{j=0}^{L-1} \sum_{i=0}^{L-1} x[i]x[j]x[j+1]
\nonumber \\
&= \frac{\gamma^2}{L} \left\{ \sum_{\ell=0}^{L-1} a_x^3[1,\ell]
+ \sum_{\ell=1}^{L-2} a_x^3[\ell,\ell+1]  \right\}.
\end{align}
%
Since we also observe $\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] + \sum_{\ell=1}^{L-2} \gamma a_x^3[\ell,\ell+1]$, we can form the ratio and solve for $\gamma$:
%
\begin{align}
\gamma = L \frac{(\gamma a_x^1) (\gamma a_x^2[1])}
{\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] 
	+ \sum_{\ell=1}^{L-2} \gamma a_x^3[\ell,\ell+1]}.
\end{align}
%
With $\gamma$ in hand, we can estimate $\sigma^2$:
%
\begin{align}
\sigma^2 = a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\frac{L (a^1_y)^2}{\gamma}
- (2 L - 1) (a_y^1)^2.
\end{align}

@@

\begin{corollary} \TODO{re-write}
	From the first three moments of the Poisson process model, one can recover the first three moments from the strongly-separated model.
\end{corollary}
\end{comment}

\TODO{What we want to have in this section}
\begin{enumerate}
	\item Recall the Poisson model (it's already described at the end of intro, so keep it short).
	\item Define a proper notion of $\gamma$ (careful with scaling $L$).
	\item State formulas for $a_y^q$ with $q = 1, 2, 3$ (ideally, including noise terms) to parallel what we did in the well-separated section. It would be better to do everything with autocorrelations rather than moments to keep the story uniform.
	\item (optional) Claim: under some assumptions (knowing $\gamma$? Others?), the autocorrelations of $y$ of order $q = 1, 2, 3$ under the Poisson model provide the same information about $x$ as the corresponding autocorrelations under the well-separated model. Could also be in Section 4.1.
\end{enumerate}

>>>>

In this section, we give the expressions for the autocorrelations of the observed signal $y$ under the Poisson model. In this model, each location $k$, $1 \le k \le N-L+1$ is the starting point of a Poisson-distributed number of random signals. More precisely, letting $s[k]$ be the number of signals starting at location $k$, $s[k] \overset{iid}{\sim} \Poisson(\gamma / L)$. Note that the expected total number of signals occurring in $y$ is equal to $(N - L + 1)\gamma / L$, and consequently the total lengths of all signals (including overlaps) divided by $N$ is equal to
%
\begin{align}
\frac{N - L + 1}{N} \cdot L \cdot \frac{\gamma  }{L} 
\quad \overset{N\to\infty}{\longrightarrow} \quad \gamma.
\end{align}
%
Therefore, in the large $N$ limit $\gamma$ may be interpreted as the signal density, as in the well-separated model.

We will show the following expressions for the autocorrelations of $y$, expressed in terms of the autocorrelations of $x$ and the parameters $\gamma$ and $\sigma$:
%
\begin{align}
a_y^1  &=  \gamma a_{x}^1, 
\label{eq:mean_micrograph2} \\
a_y^2[\ell] &=  (\gamma a_x^1)^2 + \gamma a_x^2[\ell] + \sigma^2 \delta[\ell], 
\label{eq:ac2_micrograph2} \\
a_y^3[\ell_1,\ell_2]  & = (\gamma a_x^1)^3 + \gamma a_x^1  
\cdot ( \gamma a_x^2[\ell_1] + \gamma a_x^2[\ell_2]
+ \gamma a_x^2[\ell_2-\ell_1]) + \gamma a_x^3[\ell_1,\ell_2] 
\nonumber \\
& + \gamma a_x^1 \sigma^2 \big(\delta[\ell_1]+\delta[\ell_2] +\delta[\ell_1-\ell_2] \big).
\label{eq:ac3_micrograph2}
\end{align}

@@

\TODO{Does the equivalence hold for $\ell\neq 0$?}
\begin{corollary}
	The autocorrelations are equivalent...
\end{corollary}

It is now easy to show how to solve for both $\gamma$ and $\sigma$ from these expressions, assuming $x$ is generic. Indeed, we observe the product
%
\begin{align}
(\gamma a_x^1) (\gamma a_x^2[1])
&=  \frac{\gamma^2}{L^2} \left( \sum_{i=0}^{L-1} x[i] \right)
\left(\sum_{j=0}^{L-2} x[j]x[j+1] \right)
\nonumber \\
&= \frac{\gamma^2}{L^2} \sum_{j=0}^{L-2} \sum_{i=0}^{L-1} x[i]x[j]x[j+1]
\nonumber \\
&= \frac{\gamma^2}{L} \left\{ \sum_{\ell=0}^{L-1} a_x^3[1,\ell]
+ \sum_{\ell=1}^{L-2} a_x^3[\ell,\ell+1]  \right\}.
\end{align}
%
Since we also observe $\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] + \sum_{\ell=1}^{L-2} \gamma a_x^3[\ell,\ell+1]$, \TODO{Is there a bias?} we can form the ratio and solve for $\gamma$:

\begin{align}
\gamma = L \frac{(\gamma a_x^1) (\gamma a_x^2[1])}
{\sum_{\ell=0}^{L-1} \gamma a_x^3[1,\ell] 
	+ \sum_{\ell=1}^{L-2} \gamma a_x^3[\ell,\ell+1]}.
\end{align}
%With $\gamma$ in hand, we can estimate $\sigma^2$:
%
\begin{align}
\sigma^2 = a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\frac{L (a^1_y)^2}{\gamma}
- (2 L - 1) (a_y^1)^2.
\end{align}

@@

\begin{corollary} \TODO{re-write}
	From the first three moments of the Poisson process model, one can recover the first three moments from the strongly-separated model.
\end{corollary}
\end{comment}
%
%\TODO{What we want to have in this section}
%\begin{enumerate}
%	\item Recall the Poisson model (it's already described at the end of intro, so keep it short).
%	\item Define a proper notion of $\gamma$ (careful with scaling $L$).
%	\item State formulas for $a_y^q$ with $q = 1, 2, 3$ (ideally, including noise terms) to parallel what we did in the well-separated section. It would be better to do everything with autocorrelations rather than moments to keep the story uniform.
%	\item (optional) Claim: under some assumptions (knowing $\gamma$? Others?), the autocorrelations of $y$ of order $q = 1, 2, 3$ under the Poisson model provide the same information about $x$ as the corresponding autocorrelations under the well-separated model. Could also be in Section 4.1.
%\end{enumerate}
%
